### 2023-03-23 resilience.tex
small up

% \textcolor{red}{Ask the question why use a set of propositions for $SAFE$ and $BAD$ rather than use subsets of the set of configurations ? }

We ask whether we can reach a state 
%which satisfies
in 
%
$SAFE$  in a reasonable amount of time whenever we reach a state 
% which satisfies
in
%
$BAD$. 
From this we formulate two resilience problems. First consider the case where the recovery time
is bound by a given natural number $k \geq 0$, i.e., the \emph{explicit resilience problem} for TS.

\problemx{$k$-Resilience problem }
{a transition system $(S,\rightarrow)$, an integer $k$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
{$\forall s' \in BAD, ~ s \rightarrow^* s' \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}

@@

% If we assume that the transition system yields infinite sequences of transitions, we can express the property to be evaluated in CTL by s |= AG(BAD → 0≤ j≤k EX j SAFE).

@@

%
%\problemx{Bounded Resilience problem}
%{a transition system $(S,\rightarrow)$, a state $s \in S$, $Safe, BAD \subseteq S$ and $Safe \cap BAD = \emptyset$.}
%{$\exists k \geq 0 ~ \forall s' \in BAD, ~ s \rightarrow^* s' \implies \exists s'' \in Safe ~ s' \rightarrow^{\leq k} s''$ ?\newline}
%

\problemx{Bounded Resilience problem}
{a transition system $(S,\rightarrow)$, a state $s \in S$, $Safe, Bad \subseteq S$ and $Safe \cap Bad = \emptyset$.}
{$\exists k \geq 0  \mathscr{S}$ is $k$-resilient. ?\newline}
%%

@@

pour $k,k'$ donnés, chercher si $Safe_{max}$ et $Bad_{max}$ ont un sens ? on peut faire l'union des Safe, Safe' et des Bad, Bad' prendre le max des k. Fixer k, chercher et calculer $Safe_{max}$ et $Bad_{max}$.

%
%\alain{dire juste que Safe = upward closed et BAD = downward closed.
%envisager les 3 autres cas. souvent BAD = upward closed (par ex l'exclusion mutuelle).
%4 resilience problem (U,D), (U,U), (D,U), (D,D)}

@@

@@

The resilience problem (resp. the $k$-resilience problem) in a transition system is to decide whether from a bad state, there exists a path (resp. a path of length smaller than $k$) that reaches a Safe state. Let us formalize these properties.

\problemx{resilience problem (RP)}
{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $Safe, Bad \subseteq S$.}
{$Bad \longrightarrow^{*} Safe$ ?\newline}
%
%\alain{il faudrait ne pas répéter 3 fois les mêmes imputs pour les 3 pbs: énoncer les 3 uniformes pbs d'un coup avec une fois l'input puis les 3 pbs pour un état s donné sans répéter non plus 3 fois les mêmes inputs}

\problemx{$k$-resilience problem (kRP)}
{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two sets $Safe, Bad \subseteq S$.}
{$Bad \longrightarrow^{\leq k} Safe$ ?\newline}

We add a third problem that decides whether there exists an $k$ such  that the system is $k$-resilient.
%
\problemx{bounded resilience problem (BRP)}
{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $Safe, Bad \subseteq S$.}
%{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
{$\exists k \geq 0$ such that $\mathscr{S}$ is %uniformely
 $k$-resilient ?\newline}

\alain{trouver les Bad et Safe maximum tels que S est resilient. est-ce vrai que si S est $(B_i,D_i)$-resilient alors S est $(\cap, \cup B_i,D_i)$-resilient ? on pourrait décider que Safe=complement de Bad et donc on aurait un seul ensemble (et son complement)}

These three reachability problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. 
So we restrict our framework to the class of infinite-state WSTS. Since most of decidable properties in WSTS are given as effective upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets Safe and Bad.
\alain{on peut penser à des ensembles Bad definis dans une logique booleennne sur les clos par le bas, haut, +...}

For instance, the well-known mutual exclusion property is often modelized in a $d$-counters machine by the property that a counter $c_{mutex}$ must be bounded by (usually) one. Then, the set $Safe =  \{c_{mutex} \leq 1\} \times \mathbb{N}^{d-1}$ is downward closed and $Bad =\{c_{mutex} \geq 2\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of Safe. In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that Bad is always downward closed and Safe is always upward closed.
%		
RP is decidable for lossy counter machines (LCM) with Safe and Bad semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}. In particular, the existing proofs of resilience use the decidability of reachability; but we wish to decide resilience for models with undecidable reachability.
%
In what follows, we will always suppose that a downward closed set $D$ is given by its finite set of ideals $I_D=\{J_1, J_2,...,J_n\}$ satisfying $D=\downarrow D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ and that an upward closed set $U$ is given by its finite set $Min_U$ of its minimal elements satisfying $U=\uparrow Min_U$.

Let $U,V$ be two upward-closed sets given by its set of minimal element and $D,E$ be two downward-closed sets given by its ideal decomposition. All the four  inclusions $U \subseteq V$,  $U \subseteq D$, $D \subseteq U$ and $D \subseteq E$ are decidable provided that the ordering is decidable and sets are recursive.
%
\alain{on peut penser à des ensembles bad definis dans une logique booleennne sur les clos par le bas, haut, +...}

\begin{proposition}\label{general}
$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
(bad,Safe)-resilient iff $bad \subseteq \pred^*(Safe)$.\\
$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
(bad,Safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
(bad,Safe)-resilient.\\
$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
(bad,Safe)-$k$-resilient iff $bad \subseteq \pred^k(Safe)$.
\end{proposition}

@@

\subsection{Case: $Safe=\uparrow Safe$ and $BAD=\uparrow BAD$.}

\begin{theorem}\label{up-up}
The three %uniform 
resilience problems are decidable for WSTS with effective predbasis, $Safe=\uparrow Safe$
and $BAD=\uparrow BAD$.
%	and bad is upward closed or downward closed.
\end{theorem}

\begin{proof}
Let us first solve if $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
(bad,Safe)-resilient. Since $Safe=\uparrow Safe$, and WSTS, $\pred^*(Safe)=\uparrow \pred^*(Safe)$, so $\pred^*(Safe)$ admits a finite basis $B_{\pred^*(Safe)}$. since WSTS  with effective predbasis, we may compute a finite basis of $\pred^*(Safe)$ with the backward coverability algorithm. 
%Then if bad is upward closed or downward closed, we may decide the inclusion $Bad \subseteq \uparrow B_{\pred^*(Safe)}$ which remais to compare, in both cases, two finite sets. 
if bad  is upward closed, $bad = \uparrow B_{bad}$ and $ \uparrow B_{bad} \subseteq \uparrow B_{\pred^*(Safe)}$ iff for every $b \in B_{bad}$, there is a $s \in B_{\pred^*(Safe)}$ such that $s \leq b$. 
\iffalse
similar reasonning for the other case.
\fi
hence the two %uniform
 resilience problems are decidable.

Let us examine the %uniform 
$k$-resilience problem (kRP). We may compute the least $n$ such that $bad \subseteq \uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$. $S$ is %uniformely 
$k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}
\mathieu{Si on a la résilience alors on a la $k$-résilience pour les $k$ plus grands que le $n$ minimal tel que $\uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$, mais il faut traiter aussi les $k$ plus petits.}
\mathieu{Si on a $\uparrow \pred^k(Safe) = \pred^k(Safe)$ on résoud comme ci-dessus.}
\mathieu{... mais pour les $k$ plus petits, on peut avoir $\uparrow \pred^k(Safe) \neq \pred^k(Safe)$.}
\mathieu{... et alors, c'est plus compliqué.}
\mathieu{L'important c'est d'avoir le $n$ minimal tel que $bad \subseteq \pred^n(Safe)$, qui n'est pas forcément tel que $\uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$.}and then kRP is decidable.
\end{proof}

%One begins to test whether $Bad \cap \pred^*(Safe) = \emptyset$ then never resilient car une fois dans bad j'y reste sans pouvoir aller dans Safe.\\

If $Bad \cap \pred^*(Safe) \neq \emptyset$, One computes the minimal elements $A$ of $Bad \cap \pred^*(Safe)$.
resilient iff $Bad \subseteq \pred^*(Safe)$  (eq $Bad \cap \pred^*(Safe) = Bad$)

$Bad \cap \pred^*(Safe) \neq Bad$ alors, on peut se limiter à $Bad' = Bad \cap \pred^*(Safe)$ et c'est (Bad',Safe)-resilient.

@@

\subsection{Case: $Safe=\uparrow Safe$ and $BAD=\downarrow BAD$.}

@@

%Transfering the abstract resilience problems into this framework,
%it is therefore reasonable to demand that both propositions, Safe and BAD, are given by 
%upward-closed or downward-closed sets.

@@

\begin{proposition}\label{down-up}
Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two finite sets: $I_{BAD}$ and $Min_{Safe}$.
The %uniform 
resilience problem (RP) and the %uniform 
bounded resilience problem (BRP) are decidable.
The %uniform
 $k$-resilience problem (kRP) is decidable if moreover $\mathscr{S}$ has the predbasis hypothesis.
\end{proposition}

\begin{proof}
Let $I_{BAD}=\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $BAD$ and $B_{Safe}=\{b_1,b_2,...,b_m\}$ be the minimal basis of $Safe$.
The %uniform 
resilience problem (RP) is equivalent to an infinite number of instances of the coverability problem : for all $x \in BAD$ does there exist an $j$ such that $b_j$ is coverable from $x$. This infinite set of coverability questions can be reduced to $n$ instances of the coverability problem in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.

@@

%
% iff there is a run in $\hat{S}$ from an ideal I to J such that  $\downarrow x \subseteq I$ and $\downarrow b_j \subseteq J$. This is a consequence of  
%Propositions 29 and 30 in \cite{BFM-icalp14} that establish a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ with its completion $\hat{S}$.
%' \in I(BAD)$.

To decide the %(uniform)
 bounded resilience problem (BRP), we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ? If all these $n \times m$ coverability questions are positive then we compute $K=max(k_a \mid a=1,..,m \times n)$ and we deduce that  $\mathscr{S}$ is %uniformely
  $K$-resilient. Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we have shown that both the %uniform 
  resilience problem (BRP) and the %uniform 
  bounded resilience problem (BRP) are decidable.

To decide the kRP, we may compare the input $k$ with $K$. If $k \geq K$, we conclude that is %uniformely 
$k$-resilient; if $k < K$ we cannot answer. if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, we may compute the least $n$ such that $ \uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$. $S$ is %uniformely
 $k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then kRP is decidable.
\end{proof}

@@

\subsection{Case: $Safe=\downarrow Safe$ and $BAD=\downarrow BAD$.}

If $Safe = \downarrow 0$ and $Bad = S$ . . .

% $Bad$ still necessarily contains $0$, and $0 \rightarrow^* 0$.

$Bad = J_1 \cup J_2 \cup \ldots \cup J_m$

$Safe = I_1 \cup I_2 \cup \ldots \cup I_n$

Question is: can we reach one of the $I_i$ from one of the $J_j$ ?

If the ideal reachability problem is decidable then of course the question here should be decidable as well.

But it may be possible to do more i.e.

If we can express $\downarrow Post^*( J_j)$ then the question become:
is $Safe \cap \cup_{j} \downarrow Post^*( J_j) \neq \emptyset$ ?

(we need to reach an element of $Safe$ but not every possible element)

Rather, the question is still
"is $Safe \cap \cup_{j}  Post^*( J_j) \neq \emptyset$ ?"

But maybe we can use $\downarrow Post^*( J_j)$ around here.

%%%%%%%%
\subsection{Case: $Safe=\downarrow Safe$ and $BAD=\uparrow BAD$.}
%
%		cas exclusion mutuelle
%
je prends les minimaux de bad et les idéaux de Safe, pas sur que ce soit decidable....$pred(Safe)$

If $Safe = \downarrow 0$ and $Bad = \uparrow s$ . . .

@@

In a reset petri-net, if $Bad = \uparrow q_0(\mu)$ for some $\mu \neq 0$,
and $Safe = \downarrow q(0)$,
then the resilience problem is basically \textcolor{red}{(is it?)} the single-state zero-reachability problem, and is undecidable.

\mathieu{Ici}

\begin{tabular}{ l c r }
   BAD/Safe & Up & Down \\
   Up & dec ? & dec ? \\
   Down & dec & ? \\
 \end{tabular}

>>>>

% \textcolor{red}{Ask the question why use a set of propositions for $\Safe$ and $\Bad$ rather than use subsets of the set of configurations ? }

We ask whether we can reach a state 
%which satisfies
in 
%
$\Safe$  in a reasonable amount of time whenever we reach a state 
% which satisfies
in
%
$\Bad$. 
From this we formulate two resilience problems. First consider the case where the recovery time
is bound by a given natural number $k \geq 0$, i.e., the \emph{explicit resilience problem} for TS.

\problemx{$k$-Resilience problem }
{a transition system $(S,\rightarrow)$, an integer $k$, a state $s \in S$, $\Safe, \Bad \subseteq S$ and $\Safe \cap \Bad = \emptyset$.}
{$\forall s' \in \Bad, ~ s \rightarrow^* s' \implies \exists s'' \in \Safe ~ s' \rightarrow^{\leq k} s''$ ?\newline}

@@

% If we assume that the transition system yields infinite sequences of transitions, we can express the property to be evaluated in CTL by s |= AG(\Bad → 0≤ j≤k EX j \Safe).

@@

%
%\problemx{Bounded Resilience problem}
%{a transition system $(S,\rightarrow)$, a state $s \in S$, $\Safe, \Bad \subseteq S$ and $\Safe \cap \Bad = \emptyset$.}
%{$\exists k \geq 0 ~ \forall s' \in \Bad, ~ s \rightarrow^* s' \implies \exists s'' \in \Safe ~ s' \rightarrow^{\leq k} s''$ ?\newline}
%

\problemx{Bounded Resilience problem}
{a transition system $(S,\rightarrow)$, a state $s \in S$, $\Safe, \Bad \subseteq S$ and $\Safe \cap \Bad = \emptyset$.}
{$\exists k \geq 0  \mathscr{S}$ is $k$-resilient. ?\newline}
%%

@@

pour $k,k'$ donnés, chercher si $\Safe_{max}$ et $\Bad_{max}$ ont un sens ? on peut faire l'union des \Safe, \Safe' et des \Bad, \Bad' prendre le max des k. Fixer k, chercher et calculer $\Safe_{max}$ et $\Bad_{max}$.

%
%\alain{dire juste que \Safe = upward closed et \Bad = downward closed.
%envisager les 3 autres cas. souvent \Bad = upward closed (par ex l'exclusion mutuelle).
%4 resilience problem (U,D), (U,U), (D,U), (D,D)}

@@

\newcommand{\Bad}{\textsf{Bad}}
\newcommand{\Safe}{\textsf{Safe}}

@@

The resilience problem (resp. the $k$-resilience problem) in a transition system is to decide whether from a bad state, there exists a path (resp. a path of length smaller than $k$) that reaches a \Safe state. Let us formalize these properties.

\problemx{resilience problem (RP)}
{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $\Safe, \Bad \subseteq S$.}
{$\Bad \longrightarrow^{*} \Safe$ ?\newline}
%
%\alain{il faudrait ne pas répéter 3 fois les mêmes imputs pour les 3 pbs: énoncer les 3 uniformes pbs d'un coup avec une fois l'input puis les 3 pbs pour un état s donné sans répéter non plus 3 fois les mêmes inputs}

\problemx{$k$-resilience problem (kRP)}
{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two sets $\Safe, \Bad \subseteq S$.}
{$\Bad \longrightarrow^{\leq k} \Safe$ ?\newline}

We add a third problem that decides whether there exists an $k$ such  that the system is $k$-resilient.
%
\problemx{bounded resilience problem (BRP)}
{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $\Safe, \Bad \subseteq S$.}
%{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
{$\exists k \geq 0$ such that $\mathscr{S}$ is %uniformely
 $k$-resilient ?\newline}

\alain{trouver les \Bad et \Safe maximum tels que S est resilient. est-ce vrai que si S est $(B_i,D_i)$-resilient alors S est $(\cap, \cup B_i,D_i)$-resilient ? on pourrait décider que \Safe=complement de \Bad et donc on aurait un seul ensemble (et son complement)}

These three reachability problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. 
So we restrict our framework to the class of infinite-state WSTS. Since most of decidable properties in WSTS are given as effective upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets \Safe and \Bad.
\alain{on peut penser à des ensembles \Bad definis dans une logique booleennne sur les clos par le bas, haut, +...}

For instance, the well-known mutual exclusion property is often modelized in a $d$-counters machine by the property that a counter $c_{mutex}$ must be bounded by (usually) one. Then, the set $\Safe =  \{c_{mutex} \leq 1\} \times \mathbb{N}^{d-1}$ is downward closed and $\Bad =\{c_{mutex} \geq 2\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of \Safe. In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that \Bad is always downward closed and \Safe is always upward closed.
%		
RP is decidable for lossy counter machines (LCM) with \Safe and \Bad semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}. In particular, the existing proofs of resilience use the decidability of reachability; but we wish to decide resilience for models with undecidable reachability.
%
In what follows, we will always suppose that a downward closed set $D$ is given by its finite set of ideals $I_D=\{J_1, J_2,...,J_n\}$ satisfying $D=\downarrow D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ and that an upward closed set $U$ is given by its finite set $Min_U$ of its minimal elements satisfying $U=\uparrow Min_U$.

Let $U,V$ be two upward-closed sets given by its set of minimal element and $D,E$ be two downward-closed sets given by its ideal decomposition. All the four  inclusions $U \subseteq V$,  $U \subseteq D$, $D \subseteq U$ and $D \subseteq E$ are decidable provided that the ordering is decidable and sets are recursive.
%
\alain{on peut penser à des ensembles \Bad definis dans une logique booleennne sur les clos par le bas, haut, +...}

\begin{proposition}\label{general}
$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
(\Bad,\Safe)-resilient iff $\Bad \subseteq \pred^*(\Safe)$.\\
$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
(\Bad,\Safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
(\Bad,\Safe)-resilient.\\
$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
(\Bad,\Safe)-$k$-resilient iff $\Bad \subseteq \pred^k(\Safe)$.
\end{proposition}

@@

\subsection{Case: $\Safe=\uparrow \Safe$ and $\Bad=\uparrow \Bad$.}

\begin{theorem}\label{up-up}
The % three %uniform 
resilience problem is decidable for WSTS with effective predbasis, $\Safe=\uparrow \Safe$
and $\Bad=\uparrow \Bad$.
%	and \Bad is upward closed or downward closed.
\end{theorem}

\begin{proof}
Let us first solve if $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
(\Bad,\Safe)-resilient. Since $\Safe=\uparrow \Safe$, and WSTS, $\pred^*(\Safe)=\uparrow \pred^*(\Safe)$, so $\pred^*(\Safe)$ admits a finite basis $B_{\pred^*(\Safe)}$. since WSTS  with effective predbasis, we may compute a finite basis of $\pred^*(\Safe)$ with the backward coverability algorithm. 
%Then if \Bad is upward closed or downward closed, we may decide the inclusion $\Bad \subseteq \uparrow B_{\pred^*(\Safe)}$ which remais to compare, in both cases, two finite sets. 
if \Bad  is upward closed, $\Bad = \uparrow B_{\Bad}$ and $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^*(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$. 
\iffalse
similar reasonning for the other case.
\fi
hence the %two
 %uniform
 resilience problem is decidable.
\end{proof}

\begin{corollary}
The bounded resilience problem is decidable for WSTS with effective predbasis, $\Safe=\uparrow \Safe$
and $\Bad=\uparrow \Bad$.
\end{corollary}

\begin{theorem}\label{up-up}
The % three %uniform 
$k$-resilience problem is decidable for WSTS with effective predbasis, $\Safe=\uparrow \Safe$
and $\Bad=\uparrow \Bad$.
%	and \Bad is upward closed or downward closed.
\end{theorem}

\begin{proof}
Let us examine the %uniform 
$k$-resilience problem (kRP). We may compute the least $n$ such that $\Bad \subseteq \uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$. $S$ is %uniformely 
$k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}
\mathieu{Si on a la résilience alors on a la $k$-résilience pour les $k$ plus grands que le $n$ minimal tel que $\uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$, mais il faut traiter aussi les $k$ plus petits.}
\mathieu{Si on a $\uparrow \pred^k(\Safe) = \pred^k(\Safe)$ on résoud comme ci-dessus.}
\mathieu{... mais pour les $k$ plus petits, on peut avoir $\uparrow \pred^k(\Safe) \neq \pred^k(\Safe)$.}
\mathieu{... et alors, c'est plus compliqué.}
\mathieu{L'important c'est d'avoir le $n$ minimal tel que $\Bad \subseteq \pred^n(\Safe)$, qui n'est pas forcément tel que $\uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$.}and then kRP is decidable.
\end{proof}

%One begins to test whether $\Bad \cap \pred^*(\Safe) = \emptyset$ then never resilient car une fois dans \Bad j'y reste sans pouvoir aller dans \Safe.\\

If $\Bad \cap \pred^*(\Safe) \neq \emptyset$, One computes the minimal elements $A$ of $\Bad \cap \pred^*(\Safe)$.
resilient iff $\Bad \subseteq \pred^*(\Safe)$  (eq $\Bad \cap \pred^*(\Safe) = \Bad$)

$\Bad \cap \pred^*(\Safe) \neq \Bad$ alors, on peut se limiter à $\Bad' = \Bad \cap \pred^*(\Safe)$ et c'est (\Bad',\Safe)-resilient.

@@

\subsection{Case: $\Safe=\uparrow \Safe$ and $\Bad=\downarrow \Bad$.}

@@

%Transfering the abstract resilience problems into this framework,
%it is therefore reasonable to demand that both propositions, \Safe and \Bad, are given by 
%upward-closed or downward-closed sets.

@@

\begin{proposition}\label{down-up}
Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two finite sets: $I_{\Bad}$ and $Min_{\Safe}$.
The %uniform 
resilience problem (RP) and the %uniform 
bounded resilience problem (BRP) are decidable.
\end{proposition}

\begin{proof}
Let $I_{\Bad}=\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\Bad$ and $B_{\Safe}=\{b_1,b_2,...,b_m\}$ be the minimal basis of $\Safe$.
The %uniform 
resilience problem (RP) is equivalent to an infinite number of instances of the coverability problem : for all $x \in \Bad$ does there exist an $j$ such that $b_j$ is coverable from $x$. This infinite set of coverability questions can be reduced to $n$ instances of the coverability problem in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.

@@

%
% iff there is a run in $\hat{S}$ from an ideal I to J such that  $\downarrow x \subseteq I$ and $\downarrow b_j \subseteq J$. This is a consequence of  
%Propositions 29 and 30 in \cite{BFM-icalp14} that establish a strong relation between the runs of a WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ with its completion $\hat{S}$.
%' \in I(\Bad)$.

To decide the %(uniform)
 bounded resilience problem (BRP), we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ? If all these $n \times m$ coverability questions are positive then we compute $K=max(k_a \mid a=1,..,m \times n)$ and we deduce that  $\mathscr{S}$ is %uniformely
  $K$-resilient. Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we have shown that both the %uniform 
  resilience problem (BRP) and the %uniform 
  bounded resilience problem (BRP) are decidable.
\end{proof}

\begin{proposition}
Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and the predbasis hypothesis, and two finite sets: $I_{\Bad}$ and $Min_{\Safe}$.
The %uniform
 $k$-resilience problem (kRP) is decidable.
 \end{proposition}

\begin{proof}
To decide the kRP, we may compare the input $k$ with $K$. If $k \geq K$, we conclude that is %uniformely 
$k$-resilient; if $k < K$ we cannot answer. if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, we may compute the least $n$ such that $ \uparrow \pred^n(\Safe)=  \uparrow \pred^*(\Safe)$. $S$ is %uniformely
 $k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then kRP is decidable.
\end{proof}

@@

\subsection{Case: $\Safe=\downarrow \Safe$ and $\Bad=\downarrow \Bad$.}

% In the case \Safe and \Bad are both downward closed, they can both be finite.

\begin{proposition}
The resilience problem is undecidable for WSTS with \textcolor{red}{more than one minimal element}, $\Safe=\downarrow \Safe$
and $\Bad=\downarrow \Bad$.
\end{proposition}

\begin{proof}
In the case of a WSTS with more than one minimal element $0_1$ and $0_2$, then the problem of whether $0_2$ is reachable from $0_1$ reduces itself to the resilience problem.  
\end{proof}

In the case of a WSTS with only one minimal element,  
% (ex: Petri Nets and their «Zero») 
then by downward closure the minimal element is present in both \Safe and \Bad and hence resilience always hold.

%%%%%%%%
\subsection{Case: $\Safe=\downarrow \Safe$ and $\Bad=\uparrow \Bad$.}
%
%		cas exclusion mutuelle
%
je prends les minimaux de $\Bad$ et les idéaux de $\Safe$, pas sur que ce soit decidable....$pred(\Safe)$

If $\Safe = \downarrow 0$ and $\Bad = \uparrow s$ . . .

@@

In a reset petri-net, if $\Bad = \uparrow \mu$ for some $\mu \neq 0$,
and $\Safe = \downarrow 0$,
then the resilience problem is basically \textcolor{red}{(is it?)} the single-state zero-reachability problem, and is undecidable.

\mathieu{So, the idea is that the question "is $0$ reachable from token assigment $\mu$" is undecidable, but here the question is slightly different because we start from $\uparrow \mu$ instead.}

\begin{tabular}{ l c r }
\hline   \Bad/\Safe & Up & Down \\ \hline
   Up & dec ? & dec ? \\ \hline
   Down & dec & ? \\ \hline
 \end{tabular}
